---
title: "GSE280465 EPICv2 methylation data"
author: "Paul Ruiz Pinto"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{GSE280465 EPICv2 methylation data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document
---

# GSE280465 EPICv2 methylation data

`GSE280465` provides access to processed adult DNA methylation data derived
from GEO accession
[GSE280465](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE280465).
The source data were generated on GPL33022, the Illumina Infinium
MethylationEPIC version 2 array.

The ExperimentHub resource is a `RangedSummarizedExperiment` with beta, M, and
CN assays for 163 adult samples from 47 individuals. Samples represent buccal,
saliva, dried blood spot, and peripheral blood mononuclear cell tissues.
Genomic ranges use hg38 coordinates.

The source publication is:

Apsley AT, Ye Q, Caspi A, Chiaro C, Etzel L, Hastings WJ, Heim CM, Kozlosky J,
Noll JG, Schreier HMC, Shenk CE, Sugden K, Shalev I. Cross-tissue comparison
of epigenetic aging clocks in humans. Aging Cell. 2025;24(4):e14451.
doi:10.1111/acel.14451.

# Accessing the resource

The data can be located in `ExperimentHub` using the GEO accession or the
package name. The loaded object can then be inspected with standard
`SummarizedExperiment` accessors.

```{r access-resource, eval=FALSE}
library(ExperimentHub)
library(GSE280465)
library(SummarizedExperiment)

eh <- ExperimentHub()
query(eh, "GSE280465")

se <- GSE280465()
se

se <- GSE280465_GPL33022_EPICv2()

dim(se)
assayNames(se)
colData(se)
rowRanges(se)
metadata(se)
```

The package metadata describe the hosted resource and its source.

```{r local-metadata}
metadataFile <- system.file("extdata", "metadata.csv", package = "GSE280465")
metadata <- utils::read.csv(metadataFile, stringsAsFactors = FALSE)
metadata[, c("Title", "Genome", "SourceType", "RDataClass", "DispatchClass")]
```

# Data availability

The serialized `RangedSummarizedExperiment` is hosted on Zenodo:

```{r zenodo-location}
paste0(metadata$Location_Prefix, metadata$RDataPath)
```

# Session information

```{r sessionInfo}
sessionInfo()
```
